---
title: "Title"
author: "Pablo Espin, Jefferson Pastuña"
date: "2025-03-26"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
usethis::git_vaccinate()

```

## Introduction

Introduction...

## Before to start

Before to start...

## Notame workflow

As a first step...

```{r echo=TRUE, message=FALSE}

# Notame package installation
#if (!requireNamespace("devtools", quietly = TRUE)) {
#  install.packages("devtools")
#}
#devtools::install_github("antonvsdata/notame", ref = "v0.3.1")

# Notame library call
library(notame)

# Dependency packages installation
install_dependencies

```

Then, a main path and a log system was added to have a record of each process executed.

```{r echo=TRUE, message=TRUE, warning=FALSE, error=FALSE}

# Main path
ppath <- "../B_caapi_in_vitro/"
# Log system
init_log(log_file = paste0(ppath, "Result/GCMS_log.txt"))

```

Next...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

data <- read_from_excel(file = "Data/BC_GCMS_data_to_notame2.xlsx",
                        sheet = 3, corner_row = 4, corner_column = "N",
                        rt_limits = c(2, 60), split_by = c("Column", "Ion Mode"))

```

Once the data is read, the next step was...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

modes <- construct_metabosets(exprs = data$exprs, 
                              pheno_data = data$pheno_data, 
                              feature_data = data$feature_data,
                              group_col = "Group")

```

## Preprocessing

The first step is...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Data extraction
mode <- modes$Rtx5_EI
# Change 0 value to NA
mode <- mark_nas(mode, value = 0)

```

Raw data inspection...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Boxplot
bp_raw <- plot_sample_boxplots(mode,
                               order_by = "Injection_order",
                               fill_by = "QC")
# PCA
pca_raw <- plot_pca(mode,
                    center = TRUE,
                    shape = "QC",
                    color = "QC")
# Package to plots visualization in a same windows
#if (!requireNamespace("devtools", quietly = TRUE)) {
#  install.packages("devtools")
#}
#devtools::install_github("thomasp85/patchwork")
library(patchwork)
# Plot
bp_raw + pca_raw

```


Then...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Low detection rate
mode <- flag_detection(mode, qc_limit = 0.7, group_limit = 6/9)

```

The next step...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Drift correction
corrected <- correct_drift(mode)
#corrected <- correct_drift(corrected)
# Flag low quality features
corrected <- flag_quality(corrected, condition = "RSD_r < 0.3 & D_ratio_r < 0.74")

```

Then we can visualize the data after drift correction.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Boxplot
corr_bp <- plot_sample_boxplots(corrected,
                                order_by = "QC",
                                fill_by = "QC")
# PCA
corr_pca <- plot_pca(corrected,
                     center = TRUE,
                     shape = "QC",
                     color = "QC")
# Plot
corr_pca + corr_bp

```

The next step...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

clustered <- cluster_features(corrected,
                              rt_window = 1/60,
                              corr_thresh = 0.95,
                              d_thresh = 0.85,
                              plotting = TRUE,
                              prefix = paste0(ppath, "Result/Cluster/")
                              )
compressed <- compress_clusters(clustered)

```

We can visualize data...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Boxplot
clust_bp <- plot_sample_boxplots(compressed,
                                 order_by = "QC",
                                 fill_by = "QC")
# PCA
clust_pca <- plot_pca(compressed,
                      center = TRUE,
                      shape = "QC",
                      color = "QC")
# Plot
clust_pca + clust_bp

```

Finally the data is...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

#save(compressed, file = paste0(ppath, "Result/Notame_GC-MS_out.RData"))

```

Here read this paper: https://doi.org/10.1007%2Fs11306-016-1030-9

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

# Impute missing values using random forest
# To clean data
#set.seed(615)
#imputed <- impute_rf(compressed)
# To all data
#imputed <- impute_rf(imputed, all_features = TRUE)
# Impute missing values using noise threshold
imputed <- impute_simple(compressed, value = 100, na_limit = 0)

```

We can inspect PCA plot after...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Boxplot
imp_bp <- plot_sample_boxplots(imputed,
                               order_by = "QC",
                               fill_by = "QC")
# PCA
imp_pca <- plot_pca(imputed,
                    center = TRUE,
                    shape = "Group",
                    color = "Group")
# Plot
imp_pca + imp_bp

```

The next step...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Internal Standar Normalization
# Feature table extraction
feat_table <- exprs(imputed)

# Internal standard data extraction
is_data <- feat_table["Rtx5_EI_194_052603269692a25_099312", ]

# Internal standard normalization
is_norm <- sweep(feat_table, 2, is_data , "/")


```

We can inspect PCA plot after...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide', fig.width = 7, fig.height = 4}

# Librerías necesarias
library(ggplot2)
library(dplyr)

# Eliminar el estándar interno antes de PCA
is_norm_noIS <- is_norm[rownames(is_norm) != "Rtx5_EI_194_052603269692a25_099312", ]

# Transponer para PCA 
is_norm_t <- t(is_norm_noIS)

# Realizar PCA
ei_pca <- prcomp(is_norm_t, scale. = TRUE)

# Crear tabla de scores con los metadatos
scores <- ei_pca$x %>%
  data.frame() %>%
  mutate(Sample_ID = rownames(.)) %>%
  left_join(imputed@phenoData@data, by = "Sample_ID")

# Graficar PCA
figure_1a <- ggplot(scores, aes(PC1, PC2, shape = Group, color = Group)) +
  geom_point(size = 3, stroke = 1, shape = 21, fill = "white") +
  geom_point(size = 3) +
  guides(x = guide_axis(title = "PC1 (20.05 %)"),
         y = guide_axis(title = "PC2 (17.78 %)")) +
  theme_classic() +
  theme(legend.text = element_text(face = "italic")) +
  theme(legend.position = c(0.890, 0.790),
        legend.background = element_rect(fill = "white", color = "black")) +
  theme(panel.grid = element_blank(),
        panel.border = element_rect(fill = "transparent")) +
  geom_vline(xintercept = 0, linetype = "longdash", colour = "gray") +
  geom_hline(yintercept = 0, linetype = "longdash", colour = "gray")

# Mostrar figura
figure_1a

```

Droping flagged features...

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

# Extract clean data
pqn_noflag <- drop_flagged(pqn_set)
# Extracting feature height table
peak_height <- exprs(pqn_noflag)
# Extracting Phenotipic data
pheno_data <- pqn_noflag@phenoData@data

```

Preparing data and transposing feature table.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

# Transposing feature height table
transp_table  <- t(peak_height)
# Centering and Scaling features
ei_pca <- prcomp(transp_table, center = TRUE, scale. = TRUE)

```

Plotting PCA results.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide', fig.width = 7, fig.height = 4}

# Library to left_join use
library(dplyr)
# PCA scores
scores <- ei_pca$x %>%                   # Get PC coordinates
  data.frame %>%                         # Convert to data frames
  mutate(Sample_ID = rownames(.)) %>%    # Create a new column with the sample names
  left_join(pheno_data )                 # Adding metadata
# PCA plot
figure_1a <- ggplot(scores,
       aes(PC1, PC2, shape = Group, color = Group)) +
  geom_point(size = 3) +
  guides(x=guide_axis(title = "PC1 (20.05 %)"),
         y=guide_axis(title = "PC2 (17.78 %)")) +
  theme_classic() +
  theme(legend.text = element_text(face="italic")) +
  theme(legend.position = c(0.890, 0.790),
        legend.background = element_rect(fill = "white", color = "black")) +
  theme(panel.grid = element_blank(), 
        panel.border = element_rect(fill= "transparent")) +
  geom_vline(xintercept = 0, linetype = "longdash", colour="gray") +
  geom_hline(yintercept = 0, linetype = "longdash", colour="gray")
figure_1a
# Save plot
#ggsave('Result/notame_Result/HS_GCMS/figure_1a.pdf', width = 7, height = 4, device='pdf', dpi="print")

```

Plotting loading results.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

loadings <- ei_pca$rotation %>%           # Extract loadings
  data.frame(Feature_ID = rownames(.))    # New column with feat name

```

Creating an artificial table with Feature name and Compound column.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide', fig.width = 14, fig.height = 7}

# Extracting feature identified
metab_data <- pqn_noflag[!is.na(pqn_noflag@featureData@data$Metabolite),]
# Extracting metabolite table
meta_table <- metab_data@featureData@data
# Creating a new small table of the annotated compounds
ei_compouds <- left_join(meta_table, loadings)
# Plotting results
figure_1c <- ggplot(loadings, aes(PC1, PC2)) + 
  geom_point(alpha = 0.2) +
  theme_classic() + 
  geom_point(data = ei_compouds,
             aes(shape = meta_table$IL,
                 color = meta_table$IL),
             size = 2) +
  labs(shape = 'Identification level',
       color = 'Identification level') +
  ggrepel::geom_label_repel(data = ei_compouds,
                            aes(label = meta_table$Metabolite),
                            box.padding = 0.37,
                            label.padding = 0.22,
                            label.r = 0.30,
                            cex = 2.5,
                            max.overlaps = 50,
                            min.segment.length = 0) +
  guides(x=guide_axis(title = "PC1 (20.05 %)"),
         y=guide_axis(title = "PC2 (17.78 %)")) +
  theme(legend.position = c(0.945, 0.925),
        legend.background = element_rect(fill = "white", color = "black")) +
  theme(panel.grid = element_blank(), 
        panel.border = element_rect(fill= "transparent")) +
  geom_vline(xintercept = 0, linetype = "longdash", colour="gray") +
  geom_hline(yintercept = 0, linetype = "longdash", colour="gray") +
  ggsci::scale_color_aaas()
figure_1c
# Save plot
#ggsave('Result/notame_Result/HS_GCMS/figure_1c.pdf', width = 14, height = 7, device='pdf', dpi="print")

```

# Heat map plot

ComplexHeatmap package and dependency installation.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

# ComplexHeatmap package installation
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("ComplexHeatmap")
library(ComplexHeatmap)

# ColorRamp2 package installation
#if (!requireNamespace("devtools", quietly = TRUE)) {
#  install.packages("devtools")
#}
#devtools::install_github("jokergoo/colorRamp2")
library(colorRamp2)

# Cowplot package installation
#install.packages("cowplot")
library(cowplot)

# mdatools package installation
#install_github('svkucheryavski/mdatools')
library(mdatools)

# ClassyFire package installation
#remotes::install_github('aberHRML/classyfireR')
library(classyfireR)

```

The metabolites were classified using the ClassyFireR to add these metabolite classifications to the heatmap plot.


```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

# InChI key of the metabolites you want to classify
InChI_Keys <- c('2-Heptanone' = "CATSNJVOTSVZJV-UHFFFAOYSA-N")
# Get classification
Classification_List <- purrr::map(InChI_Keys, get_classification)
Classification_List

```

Extracting and loaded of identified metabolites abundance.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Drop QC
hm_no_qc <- drop_qcs(pqn_noflag)
# Scaling by autoscaling method
hm_scl <- scale(t(exprs(hm_no_qc)), center = TRUE, scale = TRUE)
hm_scl <- t(hm_scl)
# Adding autoscaling data to notame MetaboSet
hm_scl_set <- hm_no_qc
exprs(hm_scl_set) <- hm_scl
# Extracting identified metabolite data
raw_hm <- hm_scl_set[!is.na(hm_scl_set@featureData@data$Metabolite),]
# Extracting feature height table
hm_height <- exprs(raw_hm)
# Extracting sample information
hm_pdata <- raw_hm@phenoData@data
# Extracting feature information
hm_fdata <- raw_hm@featureData@data

```

Scaling, row and top heatmap anotation.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

set.seed(1540)
# Adding row and column names
hm_scl <- hm_height
rownames(hm_scl) <- hm_fdata$Metabolite
colnames(hm_scl) <- hm_pdata$Species
# Metabolite class color
cols_metclass <- c("Benzenoids" = "#800000FF",
                   "Hydrocarbons" = "#FFA319FF",
                   "Lipids and lipid-like molecules" = "#8A9045FF",
                   "Organic oxygen compounds" = "#8DD3C7",
                   "Organohalogen compounds" = "#BEBADA",
                   "Organoheterocyclic compounds" = "#FFFFB3")
# Add row anotation to HeatMap
hm_row_ann <- rowAnnotation(`Class` = hm_fdata$classyfireR_Superclass,
                            col = list(`Class` = cols_metclass),
                            show_annotation_name = T,
                            show_legend = F)
# Species color
cols_species <- c("T. angustula" = "#E76BF3",
                  "M. fasciculata" = "#e6550d",
                  "M. fuscopilosa" = "#91cf60")
# Add top anotation to HeatMap
top_info_ann <- HeatmapAnnotation(`Species` = hm_pdata$Species,
                                  col = list(`Species` = cols_species),
                                  show_annotation_name = T,
                                  show_legend = F,
                                  border = T)
# Color scale
mycol <- colorRamp2(c(-4, 0, 4),
                    c("blue", "white", "red"))
# Heatmap matrix plotting
hm_plot <- Heatmap(hm_scl,
                   col = mycol,
                   border_gp = grid::gpar(col = "black", lty = 0.02),
                   rect_gp = grid::gpar(col = "black", lwd = 0.75),
                   clustering_distance_columns = "euclidean",
                   clustering_method_columns = "complete",
                   top_annotation = top_info_ann,
                   column_names_gp = gpar(fontface = "italic"),
                   row_names_max_width = unit(10, "cm"),
                   right_annotation = hm_row_ann,
                   show_heatmap_legend = F,
                   row_km = 3, column_km = 2)
hm_plot

```

Adding legends to heatmap.

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

# Color scale legend
lgd1 <- Legend(col_fun = mycol,
               title = "Autoscaled abundance",
               direction = "horizontal" )

# Bees species legend
lgd2 <- Legend(labels = gt_render(c("*T. angustula*",
                                    "*M. fasciculata*",
                                    "*M. fuscopilosa*")),
               legend_gp = gpar(fill = cols_species),
               title = "Bees species", ncol = 1)
# Metabolite class Legend
lgd3 <- Legend(labels = c(unique(hm_fdata$classyfireR_Superclass)) ,
               legend_gp = gpar(fill = cols_metclass), 
               title = "Metabolite superclass", ncol = 2)

```

ComplexHeatmap plot

```{r echo=TRUE, message=FALSE, warning=FALSE, error=FALSE}

set.seed(1540)
# Converting to ggplot
gg_heatmap <- grid.grabExpr(draw(hm_plot))
gg_heatmap <- ggpubr::as_ggplot(gg_heatmap)
# Legends
all_legends <- packLegend(lgd1, lgd2, lgd3, direction = "horizontal")
gg_legend <- grid.grabExpr(draw(all_legends))
gg_legend_fn <- ggpubr::as_ggplot(gg_legend)
# Heatmap plot
gcms_hm <- plot_grid(gg_legend_fn,
                     gg_heatmap, ncol = 1,
                     rel_heights = c(0.055, 0.880))
gcms_hm
# Save heatmap plot
#ggsave(filename = "Result/GCMS/GCMS_Heatmap_to_DQ3.pdf", plot = gcms_hm,
#       width = 8, height = 9, units = "in", dpi = 300, scale = 1.7)

```

Finish a record.

```{r}

finish_log()

```



